### 2025-01-18 manuscript.tex
GitHub trade bot tested and results included.

\cleardoublepage
\begingroup
\makeatletter
\let\ps@plain\ps@empty
\begin{minipage}[c][\textheight][c]{\textwidth}
	{\centering
	{\large UNIVERZITA J. SELYEHO\\SELYE JÁNOS EGYETEM}\\
	\vskip0.5cm
	{\ NÁZOV FAKULTY\\Fakulta ekonómie a informatiky\\Gazdaságtudományi és Informatikai Kar}\\
	\vfill
	{\Large NÁZOV PRÁCE\\Analýza údajov o cenách na burze v reálnom čase pomocou neurónových sietí }\\
	\vfill
	\thispagestyle{empty}
	\begin{tabular}{ll}
		Študijný program:    & Aplikovaná informatika \\
		Tanulmányi program:  & Alkalmazott Informatika\\
		Študijný odbor:      & Informatika\\
		Tanulmányi szak:     & Informatika\\
		Školiteľ:            & László Marák, PhD.\\
		Témavezető:          & László Marák, PhD.\\
		Školiace pracovisko: & Katedra informatiky\\
		Tanszék megnevezése: & Informatikai Tanszék\\
	\end{tabular}
	\vfill
	Označenie typu práce - Diplomamunka\\
	Bc. Eugen Fekete\\
    \ISBN \\
	\hfill\the\year{}, Komárno\hfill
	}
\end{minipage}
\endgroup
{
\hspace*{-2cm}
%\includegraphics[keepaspectratio, width=17cm]{./zadanie-zp_17422.pdf}
Ide jön az aláírt témakiírás
\pagebreak
}
\tableofcontents %% opcionális
\pagebreak
\listoffigures  %% opcionális
\addcontentsline{toc}{section}{Ábrák jegyzéke}

@@

			\subsubsection{Reinforcement learning}
			In reinforcement learning, a virtual object called an agent can make observation of its environment, commit actions resulting in rewards or penalties (usually defined as negative rewards) based on the outcome. \cite{HMLSKT}\\ 
			In reinforcement learning, the goal is to figure out a strategy (called a policy) that tells the agent what actions to take in different situations, in order to get the most reward over time. The policy can either be deterministic, where the the agent always picks the same action for a given situation, or stochastic, where the action is chosen probabilistically. The agent needs to learn how actions and rewards are connected, which means it needs to explore the policy space through trial and error. Policy space refers to the set of all possible policies that an agent can adopt. \cite{RLRAS}\\
			More precisely, this policy could be a neural network, which takes observations as inputs and outputs the resulting action. Alternatively, the policy could be a genetic algorithm, where each generation consists of multiple policies and new generations of policy offsprings are created based on the genetic algorithm used. Reinforcement learning is commonly used in robotics and smart home appliances like thermostats. We can also find financial securities trading support systems using reinforcement learning. \cite{HMLSKT}\\
			\textbf{Stock Trading Bot using Deep Q-Learning} \cite{STBDQ}\\
			A good example of a common approach in algorithmic trading can be illustrated by the "Stock Trading Bot using Deep Q-Learning" project created by Prabhsimran Singh. This project demonstrates an automatic stock trading bot using Deep Q-Learning, a type of reinforcement learning algorithm utilizing a deep neural network to approximate the optimal action-value function (also known as Q-function), which determines the best action to take in a given state. \cite{HCTDRL}. At any moment, the agent observes its current state (an n-day window of stock prices, which shows the stock prices of the last n days), selects and performs a buy/hold/sell action. The next state is evaluated and the agent receives a reward accordingly, based on changes in the portfolio value. The agent then updates its parameters by minimizing the computed loss. In reinforcement learning loss refers to the differnece between the target Q-value and the predicted Q-value, with the Q-value defining the optimal state-action (quality) values \cite{HMLSKT}. \\
			TEST RESULTS FROM THE CODE COMES HERE.\\
			\textbf{A Deep Reinforcement Learning Framework for the
			Financial Portfolio Management Problem} \cite{ADRLFFPMP}\\
			Another example of the use of reinforcement learning in the financial world is presented in the paper of Zhengyao Jiang, Dixing Xu and Jinjun Liang, titled "A Deep Reinforcement Learning Framework for the
			Financial Portfolio Management Problem". The paper describes the development of a system capable of automating portfolio management. Portfolio management is defined as the action of continuous reallocation of a capital into a number of financial assets. The core of the system is the Ensemble of Identical Independent Evaluators (EIIE) topology.  In an EIIE, there are multiple evaluators (models or algorithms), following the same structure or design, but each working independently examining an asset's historical data to predict future growth. The outputs of their individual evaluations are aggregated to make a final decision. The resulting scores from this evaluation are used to determine the portfolio weights for the next trading period. These weights represent the actions the agent will take. If an asset's weight increases, it'll be bought or if it goes down, it'll be sold. The portfolio weights from the last period are fed back into the EIIE for future predictions.\\
			Three different types of evaluators were tested in this work, a Convolutional Neural Network (CNN), a Recurrent Neural Network (RNN) and a Long Short Term Memory (LSTM) architecture. The framework was evaluated on a cryptocurrency exchange market in three different time period (to ensure that the tests cover different market conditions, like fast- or slow-moving market), each lasting 30 minutes. The tests included benchmarks, such as Best Stock (selects and invests in a single asset that achieved the highest final adjusted portfolio value), Uniform Buy and Hold (divides the initial capital equally among assets and holds them without rebalancing or trading), Uniform Constant
			Rebalanced Portfolio (divides the initial capital equally among assets and constantly rebalances the portfolio). Additionally, well-known strategies, such as ”Follow-the-Winner”, ”Follow-the-Loser”, ”Pattern-Matching”, and ”Meta-Learning” were also tested.\\
			At the beginning of each period, the trading agent reallocates the fund among the assets. The trading fee for certain actions on the site is calculated as the maximum possible percentage, which is 0.25\%. Theframework was tested on historical data only with ideal conditions: zero slippage (meaning the trade is executed immediately when an order is placed) and zero market impact (so the value of the trade is too insignificant to influence the market). For the portfolio preallocation the 11 most-volumed assets were selected. For training, historic price data from the past two days were used. The optimal policy was obtained using gradient ascent algorithm.\\
			\textbf{Algorithms}\\
			CONTINUE: EXPLAIN COMMON ALGORITHMS MORE IN DEPTH\\

>>>>

\cleardoublepage
\begingroup
\makeatletter
\let\ps@plain\ps@empty
\begin{minipage}[c][\textheight][c]{\textwidth}
	{\centering
	{\large UNIVERZITA J. SELYEHO\\SELYE JÁNOS EGYETEM}\\
	\vskip0.5cm
	{\ NÁZOV FAKULTY\\Fakulta ekonómie a informatiky\\Gazdaságtudományi és Informatikai Kar}\\
	\vfill
	{\Large NÁZOV PRÁCE\\Analýza údajov o cenách na burze v reálnom čase pomocou neurónových sietí }\\
	\vfill
	\thispagestyle{empty}
	\begin{tabular}{ll}
		Študijný program:    & Aplikovaná informatika \\
		Tanulmányi program:  & Alkalmazott Informatika\\
		Študijný odbor:      & Informatika\\
		Tanulmányi szak:     & Informatika\\
		Školiteľ:            & László Marák, PhD.\\
		Témavezető:          & László Marák, PhD.\\
		Školiace pracovisko: & Katedra informatiky\\
		Tanszék megnevezése: & Informatikai Tanszék\\
	\end{tabular}
	\vfill
	Označenie typu práce - Diplomamunka\\
	Bc. Eugen Fekete\\
    \ISBN \\
	\hfill\the\year{}, Komárno\hfill
	}
\end{minipage}
\endgroup
{
\hspace*{-2cm}
\pagebreak

Ide jön az aláírt témakiírás
\pagebreak
}

\chapter*{Statutory Declaration}
TEMPLATE\\
I hereby declare that this thesis has been composed by myself autonomously and that no means other than those declared
were used. In every single case, I have marked parts that were taken out of published or unpublished work, either verbatim or in a paraphrased manner, as such through a quotation.
\pagebreak

\chapter*{Acknowledgement}
TEMPLATE\\
Thank you everyone.
\pagebreak

\tableofcontents
\pagebreak

\listoffigures
\addcontentsline{toc}{section}{Ábrák jegyzéke}

@@

			\subsubsection{Reinforcement learning}
			In reinforcement learning, a virtual object called an agent can make observation of its environment, commit actions resulting in rewards or penalties (usually defined as negative rewards) based on the outcome. \cite{HMLSKT}\\ 
			In reinforcement learning, the goal is to figure out a strategy (called a policy) that tells the agent what actions to take in different situations, in order to get the most reward over time. The policy can either be deterministic, where the the agent always picks the same action for a given situation, or stochastic, where the action is chosen probabilistically. The agent needs to learn how actions and rewards are connected, which means it needs to explore the policy space through trial and error. Policy space refers to the set of all possible policies that an agent can adopt. \cite{RLRAS}\\
			More precisely, this policy could be a neural network, which takes observations as inputs and outputs the resulting action. Alternatively, the policy could be a genetic algorithm, where each generation consists of multiple policies and new generations of policy offsprings are created based on the genetic algorithm used. Reinforcement learning is commonly used in robotics and smart home appliances like thermostats. We can also find financial securities trading support systems using reinforcement learning. \cite{HMLSKT}\\
			\textbf{Stock Trading Bot using Deep Q-Learning} \cite{STBDQ}\\
			A good example of a common approach in algorithmic trading can be illustrated by the "Stock Trading Bot using Deep Q-Learning" project created by Prabhsimran Singh. This project demonstrates an automatic stock trading bot using Deep Q-Learning, a type of reinforcement learning algorithm utilizing a deep neural network to approximate the optimal action-value function (also known as Q-function), which determines the best action to take in a given state. \cite{HCTDRL}. At any moment, the agent observes its current state (an n-day window of stock prices, which shows the stock prices of the last n days), selects and performs a buy/hold/sell action. The next state is evaluated and the agent receives a reward accordingly, based on changes in the portfolio value. The agent then updates its parameters by minimizing the computed loss. In reinforcement learning loss refers to the differnece between the target Q-value and the predicted Q-value, with the Q-value defining the optimal state-action (quality) values \cite{HMLSKT}. \\
			Since the project is open-source and published under the MIT license, anyone can analyze and run the code. The pretrained models included in the repository were trained on Google stock using historical data from 2010. For personal testing, the provided historical Google stock data from 2019 was used.\\
			The simulation starts with a portfolio worth \$1050 in January. By late December, the portfolio's value increased to \$1359.34, generating a gross profit of \$309.34. This is a gross profit because the simulation does not account for fees required to execute the trades. The evolution of the portfolio's value is visualized using a script included in the repository.\\
			\begin{figure}[H]
			\begin{center}
				\includegraphics[width=\linewidth]{kep/trade_bot_visualization.png}
				\caption{The evolution of portfolio value using a pretrained trading bot.}
			\end{center}
			\end{figure}
			Looking at Figure 1.1, we can see that the algorithm tends to sell when the price increases and buy when the price decreases.\\

			\textbf{A Deep Reinforcement Learning Framework for the
			Financial Portfolio Management Problem} \cite{ADRLFFPMP}\\
			Another example of the use of reinforcement learning in the financial world is presented in the paper of Zhengyao Jiang, Dixing Xu and Jinjun Liang, titled "A Deep Reinforcement Learning Framework for the
			Financial Portfolio Management Problem". The paper describes the development of a system capable of automating portfolio management. Portfolio management is defined as the action of continuous reallocation of a capital into a number of financial assets. The core of the system is the Ensemble of Identical Independent Evaluators (EIIE) topology.  In an EIIE, there are multiple evaluators (models or algorithms), following the same structure or design, but each working independently examining an asset's historical data to predict future growth. The outputs of their individual evaluations are aggregated to make a final decision. The resulting scores from this evaluation are used to determine the portfolio weights for the next trading period. These weights represent the actions the agent will take. If an asset's weight increases, it'll be bought or if it goes down, it'll be sold. The portfolio weights from the last period are fed back into the EIIE for future predictions.\\
			Three different types of evaluators were tested in this work, a Convolutional Neural Network (CNN), a Recurrent Neural Network (RNN) and a Long Short Term Memory (LSTM) architecture. The framework was evaluated on a cryptocurrency exchange market in three different time period (to ensure that the tests cover different market conditions, like fast- or slow-moving market), each lasting 30 minutes. The tests included benchmarks, such as Best Stock (selects and invests in a single asset that achieved the highest final adjusted portfolio value), Uniform Buy and Hold (divides the initial capital equally among assets and holds them without rebalancing or trading), Uniform Constant
			Rebalanced Portfolio (divides the initial capital equally among assets and constantly rebalances the portfolio). Additionally, well-known strategies, such as ”Follow-the-Winner”, ”Follow-the-Loser”, ”Pattern-Matching”, and ”Meta-Learning” were also tested.\\
			At the beginning of each period, the trading agent reallocates the fund among the assets. The trading fee for certain actions on the site is calculated as the maximum possible percentage, which is 0.25\%. Theframework was tested on historical data only with ideal conditions: zero slippage (meaning the trade is executed immediately when an order is placed) and zero market impact (so the value of the trade is too insignificant to influence the market). For the portfolio preallocation the 11 most-volumed assets were selected. For training, historic price data from the past two days were used. The optimal policy was obtained using gradient ascent algorithm.\\
			\textbf{Algorithms}\\
			CONTINUE: EXPLAIN COMMON ALGORITHMS MORE IN DEPTH\\
